---
title: "Homework 3"
author: Emily Bamforth
date: 10 October 2020
output: github_document
---

```{r set_up}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Load in the data.

```{r load_data}
data("instacart")
```

This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Each observation in the dataset is based on the `order_id`. There are 15 variables in this dataset, which tell us about the order, the user, and the items, including `order_id`, `user_id`, `product_name`, and `order_hour_of_day`. Location of items are provided by department, aisle and item name.

Information about aisles: how many are there, and from which aisles are the most items purchased?

```{r aisles}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles in this dataset. The most items are purchased from the `fresh vegetables` aisle (150609 items) and the `fresh fruits` aisle (150473 items).

Here's a plot about these aisles:

```{r aisle_plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

A table about popular items:

```{r popular_table}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

And a table about the mean hour of each day of the week when Pink Lady Apples and Coffee Ice Cream are ordered:

```{r apples_icecream_table}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
pivot_wider(
  names_from = order_dow,
  values_from = mean_hour) %>% 
  knitr::kable()
```

## Problem 2

First, read the data.

```{r accel_df}
accel_df =
  read_csv("data/accel_data.csv")
```

Let's tidy the data.

```{r tidy}
accel_tidy_df =
  pivot_longer(
    accel_df,
    activity.1:activity.1440,
    names_to = "minute_of_day",
    values_to = "activity_amount"
  ) %>% 
  mutate(minute_of_day = gsub('activity.', '', minute_of_day)) %>%
  mutate(minute_of_day = as.numeric(minute_of_day)) %>% 
  mutate(day = as.factor(day),
         day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>% 
  mutate(weekend_weekday = case_when(
day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
day %in% c("Saturday","Sunday") ~ "weekend")
  )
```

This dataset consists of data from an accelerometer, tracking the activity of a 63-year-old male who was diagnosed with congestive heart failure. The dataset has `r nrow(accel_tidy_df)` rows and `r ncol(accel_tidy_df)` columns. We can see the `activity_amount` for each `minute_of_day` across a five-week period. Each observation is a minute during this time frame, and there are `r nrow(accel_tidy_df)` such observations. Other variables include the `week` number (out of 5 weeks), the `day` of the week, and a variable that tells us whether the activity took place on the weekend or weekday.
